---
title: "Medium2"
author: "Avinash Barnwal"
date: "3/24/2019"
output: html_document
---

```{r setup,include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import library and Data

I have used Lung dataset with sex being the only predictor. 

```{r Lib}
library(xgboost)
library("survival")
data("lung")
lung[,"y"] = ifelse(lung[,"status"]==1,-lung[,'time'],lung[,'time'])
```

## Results

```{r code,echo=T}
param <- list(  objective   = "survival:cox",
                eta         = 0.01,
                max.depth   = 2,
                subsample   = 0.5,
                nthread     = 2
)


df_train  <- xgb.DMatrix(as.matrix(lung[,"sex"]), label = as.matrix(lung[,"y"]))
bstSparse <- xgb.cv(data = df_train, nrounds = 10, params = param, nfold = 5, showsd = F,prediction = T)

```


Sample output in console

Iter Number : 94 Prediction: -0.682183 True: 269  
Iter Number : 95 Prediction: -0.682183 True: 270  
Iter Number : 96 Prediction: -0.682183 True: -279  
Iter Number : 97 Prediction: -0.682183 True: 284  
Iter Number : 98 Prediction: -0.682183 True: 285  
Iter Number : 99 Prediction: -0.703774 True: 285  
Iter Number : 100 Prediction: -0.682183 True: 286  
Iter Number : 101 Prediction: -0.682183 True: 288  
Iter Number : 102 Prediction: -0.682183 True: 291  
Iter Number : 103 Prediction: -0.703774 True: -292  
Iter Number : 104 Prediction: -0.682183 True: -292  
